Illustrative trajectories

### 2025-03-17 R/simulation_study/examples/illustrative_trajectories.Rmd
try some new smooth functions omega

#get functions to simulate trajectories
source(file.path(here("R","simulation_study_CRCVM","CVM_functions.R")))

@@

This type of behavior can be obtained with hyperbolic tangent function.

@@

fomega_tanh=function(cov_data,D0=0.5,omega0=60*pi/2,lambda=2,kappa=0.2) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$ExpShore
  }
  coeff=exp(-kappa*(Dshore/D0)^2)
  omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
  return(omega)
}

@@

#grid 
Dshore=seq(from=0.05,to=4,length.out=30)
theta=seq(from=-pi+0.1,to=pi-0.1,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")

#smooth omega
omega_tanh=matrix(fomega_tanh(grid),30,30)

# dynamic plots for omega

plot_omega_tanh <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(omega_tanh), end = max(omega_tanh), size = 5,color="black")),
                           x = ~theta,y = ~Dshore,z=~omega_tanh,
                           colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
  layout(title=paste("Omega"),
         scene=list(xaxis = list(title = "Theta",showgrid = F),
                    yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_omega_tanh
```

@@

  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }

@@

@@

```{r plot omega sqrt}
#smooth omega
omega_sqrt=matrix(fomega_sqrt(grid),30,30)

# dynamic plots for omega

plot_omega_sqrt <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(omega_sqrt), end = max(omega_sqrt), size = 5,color="black")),
                           x = ~theta,y = ~Dshore,z=~omega_sqrt,
                           colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
  layout(title=paste("Omega"),
         scene=list(xaxis = list(title = "Theta",showgrid = F),
                    yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_omega_sqrt
```

These smooth functions $\omega$ can produce a SDE model that is constrained within a restricted domain of $\R^2$.
The trajectories avoid the shoreline, but are not attracted to it. To model a tendecy to follow the shoreline, we need to modify a bit the smooth function to get a rotation of the velocity toward the shore when it is on the verge of leaving it. We can do this in several ways. 
First, we use a sin function.

```{r define and plot omega sin}
# Define the extended function f(theta)
fomega_sin<- function(cov_data, a =4*pi,b=80*pi,D0=0.8) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }
  ifelse(abs(theta) < pi/2, 
         -a * sin(2*theta)*exp(-(Dshore/D0)^2),
         ifelse(theta>0,
               sign(theta)*b*abs(theta-pi/2)^2*exp(-(Dshore/D0)^2),
               sign(theta)*b*abs(theta+pi/2)^2*exp(-(Dshore/D0)^2)))
}

#smooth omega
omega_sin=matrix(fomega_sin(grid),30,30)

# dynamic plots for omega

plot_omega_sin <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(omega_sin), end = max(omega_sin), size = 5,color="black")),
                           x = ~theta,y = ~Dshore,z=~omega_sin,
                           colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
  layout(title=paste("Omega"),
         scene=list(xaxis = list(title = "Theta",showgrid = F),
                    yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_omega_sin

@@

Now we use a cubic polynomial with an additional gaussian kernel.

```{r}
fomega_cubic=function(cov_data,a=1,b=5,D0=0.3,D1=0.8,sigma_theta=pi/8,sigma_D=0.3){
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }

  a*theta*(theta-pi/2)*(theta+pi/2)*exp(-Dshore/D0)/Dshore+
      b*(exp(-1/2*(theta+pi/2/sqrt(3))^2/sigma_theta^2)-
           exp(-1/2*(theta-pi/2/sqrt(3))^2/(sigma_theta)^2))*exp(-1/2*(Dshore-D1)^2/sigma_D^2)
}

#smooth omega
omega_cubic=matrix(fomega_cubic(grid),30,30)

# dynamic plots for omega

plot_omega_cubic <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(omega_cubic), end = max(omega_cubic), size = 5,color="black")),
                           x = ~theta,y = ~Dshore,z=~omega_cubic,
                           colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
  layout(title=paste("Omega"),
         scene=list(xaxis = list(title = "Theta",showgrid = F),
                    yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_omega_cubic
```

@@

fomega_minimal=function(cov_data,nu=4,D0=0.5,D1=1.5,a=1,b=10,sigma_D=0.5,sigma_theta=pi/8) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  ifelse(abs(theta)>pi/2,a*sign(theta)*(pi/2-abs(theta))*nu*cos(theta)/Dshore*exp(-Dshore/D0),-sign(theta)*b*
    exp(-1/2*(theta-sign(theta)*pi/4)^2/sigma_theta^2)*exp(-Dshore/D1))
}

#smooth omega
omega_minimal=matrix(fomega_minimal(grid),30,30)

# dynamic plots for omega

plot_omega_minimal <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(omega_minimal), end = max(omega_minimal), size = 5,color="black")),
                           x = ~theta,y = ~Dshore,z=~omega_minimal,
                           colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
  layout(title=paste("Omega"),
         scene=list(xaxis = list(title = "Theta",showgrid = F),
                    yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_omega_minimal
```

We can approximate these surfaces using tensor splines

```{r approximate tanh omega with splines}
# Splines smooth for omega -------------------- 
n <- 100000
SP_DF=5

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.2,8);
samples=data.frame(theta=theta,DistanceShore=DistanceShore)

#define grid of values
theta_v <- seq(-pi,pi,length=30)
Dshore_v<- seq(0.2,8,length=30)
pr <- data.frame(theta=rep(theta_v,30),DistanceShore=rep(Dshore_v,rep(30,30)))

#points on the surface perturbed by gaussian noise
f <- fomega_tanh(samples)
y <- f+0.1*rnorm(n)

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=SP_DF),knots=list(theta=seq(-pi,pi,len=SP_DF),
                                                       DistanceShore=seq(0.2,8,len=SP_DF)))

ExpShore=1/DistanceShore

#fit with bivariate splines te
m2 <- gam(y~te(theta,ExpShore,k=SP_DF),knots=list(theta=seq(-pi,pi,len=SP_DF),
                                                  ExpShore=seq(1/8,1/0.2,len=SP_DF)))

fomega_tanh_splines=function(cov_data) {

  if (is.null(cov_data$ExpShore)){
    omega=predict(m1,newdata=cov_data)
  }
  else if (is.null(cov_data$DistanceShore)){
    omega=predict(m2,newdata=cov_data)
  }

  return(as.numeric(omega))
}

```{r plot tanh parametric and spline omega}

@@

# Grid 
Dshore = seq(from = 0, to = 2, length.out = 30)
theta = seq(from = -pi + 0.1, to = pi - 0.1, length.out = 30)
grid <- as.data.frame(expand.grid(Dshore, theta))
colnames(grid) = c("DistanceShore", "theta")

# Smooth omega
z = matrix(fomega_cubic_advanced(grid), 30, 30)

@@

data_sim_rect_tanh=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

@@

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega_tanh,fnu=fnu_constant,v0=v0,x0=x0[i,],
                    times=times,land=rect_border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  cbind(data_sim,data_shore)
}

@@

@@

#subsample to 10 minutes
data_sim_rect_tanh_sub=data_sim_rect_tanh[seq(1,length(data_sim_rect_tanh$y1),by=60),]

ggplot() +
    geom_sf(data = rect_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_rect_tanh_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_rect_tanh_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

The process is avoiding land by rotating when it approaches the boundary. 
Now, to have trajectories that tends to remain close to the boundary and move along it we can use the function \texttt{fomega_cubic} instead.

fomega_cubic_rect_attractive=function(cov_data) {

  fomega_cubic(cov_data,a=3,D0=0.5,D1=1.5,sigma_theta=pi/8,sigma_D=0.5,b=5)

@@

data_sim_rect_cubic_attractive=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega_cubic_rect_attractive,fnu=fnu_constant,v0=v0,x0=x0[i,],
                    times=times,land=rect_border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  cbind(data_sim,data_shore)
}

@@

#subsample to 10 minutes
data_sim_rect_cubic_attractive_sub=data_sim_rect_cubic_attractive[seq(1,length(data_sim_rect_cubic_attractive$y1),by=60),]

ggplot() +
    geom_sf(data = rect_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_rect_cubic_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_rect_cubic_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

```{r simulate with omega cubic repulsive in rectangle,cache=TRUE}

fomega_cubic_rect_repulsive=function(cov_data) {

  fomega_cubic(cov_data,a=2,D0=0.5,D1=1.5,sigma_theta=pi/10,sigma_D=0.5,b=0)

@@

data_sim_rect_cubic_repulsive=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega_cubic_rect_repulsive,fnu=fnu_constant,v0=v0,x0=x0[i,],
                    times=times,land=rect_border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  cbind(data_sim,data_shore)
}

@@

#subsample to 10 minutes
data_sim_rect_cubic_repulsive_sub=data_sim_rect_cubic_repulsive[seq(1,length(data_sim_rect_cubic_repulsive$y1),by=60),]

ggplot() +
    geom_sf(data = rect_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_rect_cubic_repulsive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_rect_cubic_repulsive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

```{r simulate with omega minimal in rectangle,cache=TRUE}

fomega_minimal_rect_attractive=function(cov_data) {

@@

data_sim_rect_minimal_attractive=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega_minimal_rect_attractive,fnu=fnu_constant,v0=v0,x0=x0[i,],
                    times=times,land=rect_border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  cbind(data_sim,data_shore)
}

@@

#subsample to 10 minutes
data_sim_rect_minimal_attractive_sub=data_sim_rect_minimal_attractive[seq(1,length(data_sim_rect_minimal_attractive$y1),by=60),]

ggplot() +
    geom_sf(data = rect_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_rect_minimal_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_rect_minimal_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

Until now, we kept the parameter $\tau$ constant. We can emphasize the tendency to move along the boundary by using a smooth $\tau$ given by the functions \texttt{ftau_tanh} or \texttt{ftau_gaussian}. Here, we plot trajectories for the SDE model where the smoth $\omega$ parameter is given by \texttt{fomega_cubic} and the smooth $\tau$ parameter is given by \texttt{ftau\_gaussian}.

```{r simulate with omega cubic tau gaussian in rectangle,cache=TRUE}

data_sim_rect_cubic_gaussian_attractive=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  res=sim_CRCVM(ftau=ftau_gaussian,fomega=fomega_cubic_rect_attractive,fnu=fnu_constant,v0=v0,x0=x0[i,],
                    times=times,land=rect_border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  cbind(data_sim,data_shore)
}

```{r}
#subsample to 10 minutes
data_sim_rect_cubic_gaussian_attractive_sub=data_sim_rect_cubic_gaussian_attractive[seq(1,length(data_sim_rect_cubic_gaussian_attractive$y1),by=60),]

ggplot() +
    geom_sf(data = rect_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_rect_cubic_gaussian_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_rect_cubic_gaussian_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

@@

data_sim_rect_hanks=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

@@

```{r}
#subsample to 10 minutes
data_sim_rect_hanks_sub=data_sim_rect_hanks[seq(1,length(data_sim_rect_hanks$time),by=60),]

ggplot() +
    geom_sf(data = rect_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_rect_hanks_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_rect_hanks_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

@@

#change parameter values 
fomega_cubic_circ_attractive=function(cov_data) {

   fomega_cubic(cov_data,a=30,D0=0.5,D1=3,sigma_theta=pi/10,sigma_D=1,b=5)
}

data_sim_circ_cubic_attractive=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega_cubic_circ_attractive,fnu=fnu_constant,v0=v0,x0=x0[i,],
                    times=times,land=circ_border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  cbind(data_sim,data_shore)
}

@@

#subsample to 10 minutes
data_sim_circ_cubic_attractive_sub=data_sim_circ_cubic_attractive[seq(1,length(data_sim_circ_cubic_attractive$y1),by=60),]

ggplot() +
    geom_sf(data = circ_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_circ_cubic_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_circ_cubic_attractive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

@@

#change parameter values 
fomega_cubic_circ_repulsive=function(cov_data) {

   fomega_cubic(cov_data,a=30,D0=0.5,D1=1.5,sigma_theta=pi/10,sigma_D=0.5,b=0)
}

data_sim_circ_cubic_repulsive=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega_cubic_circ_repulsive,fnu=fnu_constant,v0=v0,x0=x0[i,],
                    times=times,land=circ_border,verbose=FALSE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore[,c("p1","p2")]
  cbind(data_sim,data_shore)
}

@@

#subsample to 10 minutes
data_sim_circ_cubic_repulsive_sub=data_sim_circ_cubic_repulsive[seq(1,length(data_sim_circ_cubic_repulsive$y1),by=60),]

ggplot() +
    geom_sf(data = circ_border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sim_circ_cubic_repulsive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sim_circ_cubic_repulsive_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") 
```

@@

data_sim_circ_hanks=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

@@

```{r}
#subsample to 10 minutes
data_sim_circ_hanks_sub=data_sim_circ_hanks[seq(1,length(data_sim_circ_hanks$time),by=60),]

>>>>

#get functions to simulate trajectories
source(file.path(here("R","simulation_study","CVM_functions.R")))

@@

This type of behavior can be obtained with several function.
We start by using the hyperbolic tangent function.

@@

fomega_tanh=function(cov_data,D0=0.5,omega0=30*pi/2,lambda=2,kappa=0.2) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  coeff=exp(-kappa*(Dshore/D0)^2)
  omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
  return(omega)
}
```

get_omega_plot<-function(fomega,Dshore,theta) {

  #grid of values
  grid<- as.data.frame(expand.grid(Dshore,theta))
  colnames(grid)=c("BoundaryDistance","BoundaryAngle")

  #smooth omega
  npoints=length(Dshore)
  omega=matrix(fomega(grid),npoints,npoints)

  # dynamic plots for omega 
  plot_omega <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(omega), end = max(omega), size = 5,color="black")),
                             x = ~theta,y = ~Dshore,z=~omega,
                           colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
  layout(title=paste("Omega"),
         scene=list(xaxis = list(title = "Theta",showgrid = F),
                    yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

  return (plot_omega)
}

Dshore=seq(from=0.05,to=4,length.out=30)
theta=seq(from=-pi+0.1,to=pi-0.1,length.out=30)

get_omega_plot(fomega_tanh,Dshore,theta)

@@

@@

  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

@@

get_omega_plot(fomega_sqrt,Dshore,theta)

@@

These smooth functions $\omega$ will produce a SDE model that is constrained within a restricted domain of $\R^2$.
The trajectories avoid the shoreline, but are not attracted to it. To model a tendency to follow the shoreline, we need to modify 
a bit the smooth function to get a rotation of the velocity toward the shore when it is on the verge of leaving it. 
We can do this in several ways. First, we use a sin function.

```{r define and plot omega sin}
# Define the extended function f(theta)
fomega_sin<- function(cov_data, a =4*pi,b=10*pi,D0=0.8) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  ifelse(abs(theta) < pi/2, 
         -a * sin(2*theta)*exp(-(Dshore/D0)^2),
         ifelse(theta>0,
               sign(theta)*b*abs(theta-pi/2)^2*exp(-(Dshore/D0)^2),
               sign(theta)*b*abs(theta+pi/2)^2*exp(-(Dshore/D0)^2)))
}

get_omega_plot(fomega_sin,Dshore,theta)

@@

In these functions, there is no mid range effect of the shore on the drift of the process, which means
that when the process is far enough from the shore, it cannot be attracted back.
To resolve this, we add a gaussian kernel in the functions.

fomega_gaussian= function(cov_data,b=1,D1=1.5,theta1=pi/2/sqrt(3),sigma_D=1,sigma_theta=0.6) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  b*(exp(-1/2*(theta+theta1)^2/sigma_theta^2)-
           exp(-1/2*(theta-theta1)^2/(sigma_theta)^2))*exp(-1/2*(Dshore-D1)^2/sigma_D^2)
}

get_omega_plot(fomega_gaussian,Dshore,theta)

We define a product of sine and cosine with an additional gaussian attractive term

```{r define and plot omega sin}
# Define the extended function f(theta)
fomega_sin_cos<- function(cov_data,omega_0=10,D0=0.8,p=1,b=1,D1=1,theta1=pi/4,sigma_D=1,sigma_theta=0.5) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

  attraction <- fomega_gaussian(cov_data,b,D1,theta1,sigma_D,sigma_theta)

  if (p%%2==0) {
    repulsion <- -omega_0*sign(theta)*sin(theta/2)^p*cos(theta)*D0/Dshore*exp(-Dshore/D0)
  }
  else {
    repulsion <- -omega_0*sin(theta/2)^p*cos(theta)*D0/Dshore*exp(-Dshore/D0)
  }
  attraction+repulsion
}

get_omega_plot(fomega_sin_cos,Dshore,theta)

Now we use a polynomial with an additional gaussian kernel to model attraction toward the shore.

```{r}
fomega_poly=function(cov_data,a=0.2,b=5,D0=0.3,D1=1.5,sigma_D=1,sigma_theta=pi/8,
                     theta1=pi/2/sqrt(3),p=3) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

  if (p%%2==0) {
    repulsion<- a*sign(theta)*theta^p*(theta-pi/2)*(theta+pi/2)*D0/Dshore*exp(-Dshore/D0)
  }
  else {
    repulsion<- a*theta^p*(theta-pi/2)*(theta+pi/2)*D0/Dshore*exp(-Dshore/D0)
  }

  attraction<- fomega_gaussian(cov_data,b,D1,theta1,sigma_D,sigma_theta)

  return (repulsion+attraction)
}

get_omega_plot(fomega_poly,Dshore,theta)
```
```{r}
fomega_quadratic=function(cov_data,a=0.2,D0=0.3,D1=1.5) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle

    return (a*(theta-pi/2)*(theta+pi/2)*D0/Dshore*exp(-Dshore/D0))
}

get_omega_plot(fomega_quadratic,Dshore,theta)
```

@@

fomega_minimal=function(cov_data,nu=4,D0=0.5,D1=1.5,a=1,b=10,sigma_D=0.5,sigma_theta=pi/8) {
  Dshore=cov_data$BoundaryDistance
  theta=cov_data$BoundaryAngle
  ifelse(abs(theta)>pi/2,a*sign(theta)*(pi/2-abs(theta))*nu*cos(theta)/Dshore*exp(-Dshore/D0),-sign(theta)*b*
    exp(-1/2*(theta-sign(theta)*pi/4)^2/sigma_theta^2)*exp(-Dshore/D1))
}

get_omega_plot(fomega_minimal,Dshore,theta)
```

We create and save the perpective plot for the paper.

```{r persp plot cubic parametric omega}

#Define function omega with adequate parameters
fomega_cubic<-function(cov_data) {
  fomega_poly(cov_data,a=2,b=5,D0=0.3,D1=0.8,sigma_D=0.3,sigma_theta=pi/8,
                     theta1=pi/2/sqrt(3),p=1)
}

@@

# Grid 
Dshore = seq(from = 0.05, to = 2, length.out = 30)
theta = seq(from = -pi + 0.1, to = pi - 0.1, length.out = 30)
grid <- as.data.frame(expand.grid(Dshore, theta))
colnames(grid) = c("BoundaryDistance", "BoundaryAngle")

# Smooth omega
z = matrix(fomega_cubic(grid), 30, 30)

@@

simulate_paralell<-function(ftau,fnu,fomega,v0=v0,x0=x0,times,border,n_cores=6,seed=42) {

  #start cluster
  cl <- makeCluster(n_cores) 
  registerDoParallel(cl)

  N<-nrow(x0)

  #paralell loop
  data=foreach (i=1:N,.combine='rbind',.export=c("sim_CRCVM","fomega_tanh","fomega_poly",
                "fomega_sin","fomega_sin_cos","fomega_sqrt","fomega_gaussian","fomega_minimal","fomega_quadratic"),
                .packages=c("progress","MASS","sf","mgcv","smoothSDE")) %dopar% {

@@

    res=sim_CRCVM(ftau=ftau,fomega=fomega,fnu=fnu,v0=v0,x0=x0[i,],
                      times=times,land=border,verbose=FALSE)

    data_sim=res$sim
    data_sim$ID=factor(rep(i,length(data_sim$y1)))
    data_shore=res$boundary[,c("p1","p2")]
    cbind(data_sim,data_shore)
  }

@@

  return (data)
}

data_rect_tanh<-simulate_paralell(ftau_constant,fnu_constant,fomega_tanh,v0,x0,times,rect_border)

```{r plot rect_tanh, message=FALSE}

get_trajectory_plot<-function(data,border,by=60) {

  #subsample to 10 minutes
  data_sub=data[seq(1,nrow(data),by=by),]

  #create plot
  plot<- ggplot() +
    geom_sf(data = border$geometry, fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_sub, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none")

  return (plot)
}

get_trajectory_plot(data_rect_tanh,rect_border)
```

The process is avoiding land by rotating when it approaches the boundary. 
Now, to have trajectories that tends to remain close to the boundary and move along it we can use the function \texttt{fomega_poly} with 
a non zero attractive gaussian term instead.

fomega_cubic_rect=function(cov_data) {

  fomega_poly(cov_data,a=1,b=5,D0=0.5,D1=1.5,sigma_D=0.5,sigma_theta=pi/8,p=1)

data_rect_cubic<-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_rect,v0,x0,times,rect_border)

@@

get_trajectory_plot(data_rect_cubic,rect_border)
```

```{r simulate with omega cubic repulsive in rectangle,cache=TRUE,message=FALSE}

fomega_cubic_repulsive_rect=function(cov_data) {

  fomega_poly(cov_data,a=1,b=0,D0=0.5,D1=1.5,sigma_D=0.5,sigma_theta=pi/8,p=1)

@@

data_rect_cubic_repulsive<-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_repulsive_rect,
                                             v0,x0,times,rect_border)

```{r}
get_trajectory_plot(data_rect_cubic_repulsive,rect_border)
```

We can also increase the parameter $p$ in the polynomial omega function to get a steeper function
at the extremities.

```{r simulate with omega cubic attractive in rectangle,cache=TRUE,message=FALSE}

fomega_poly_rect=function(cov_data) {

  fomega_poly(cov_data,a=1,b=0,D0=0.5,D1=1.5,sigma_D=0.5,sigma_theta=pi/8,p=2)

data_rect_poly<-simulate_paralell(ftau_constant,fnu_constant,fomega_poly_rect,v0,x0,times,rect_border)

@@

get_trajectory_plot(data_rect_poly,rect_border)
```

We now try our \texttt{fome_sin_cos} function defined with a product of sin and cosine.

```{r simulate with omega cubic attractive in rectangle,cache=TRUE,message=FALSE}

fomega_sin_cos_rect=function(cov_data) {

  fomega_sin_cos(cov_data,omega_0=10,D0=0.8,p=1,b=0,D1=1,theta1=pi/4,sigma_D=1,sigma_theta=0.5)

@@

data_rect_sin_cos<-simulate_paralell(ftau_constant,fnu_constant,fomega_sin_cos_rect,v0,x0,times,rect_border)

@@

get_trajectory_plot(data_rect_sin_cos,rect_border)
```

```{r simulate with omega minimal in rectangle,cache=TRUE,message=FALSE}

fomega_minimal_rect=function(cov_data) {

@@

data_rect_minimal<-simulate_paralell(ftau_constant,fnu_constant,fomega_minimal_rect,v0,x0,times,rect_border)

@@

get_trajectory_plot(data_rect_minimal,rect_border)
```

Until now, we kept the parameter $\tau$ constant. We could emphasize the tendency to move along the boundary by using a smooth $\tau$ given by the functions \texttt{ftau_tanh} or \texttt{ftau_gaussian}.

@@

data_rect_hanks=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

@@

get_trajectory_plot(data_rect_hanks,rect_border)
```

@@

#change parameter values 
fomega_cubic_attractive_circ=function(cov_data) {

   fomega_poly(cov_data,a=30,D0=0.5,D1=3,sigma_theta=pi/10,sigma_D=1,b=5,p=1,theta1=pi/2)
}

data_circ_cubic_attractive <-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_attractive_circ,v0,x0,times,circ_border)

@@

get_trajectory_plot(data_circ_cubic_attractive,circ_border)
```

@@

#change parameter values 
fomega_cubic_repulsive_circ=function(cov_data) {

   fomega_poly(cov_data,a=3,D0=0.5,D1=1.5,sigma_theta=pi/10,sigma_D=0.5,b=0,theta=pi/2,p=1)
}

data_circ_cubic_repulsive <-simulate_paralell(ftau_constant,fnu_constant,fomega_cubic_repulsive_circ,v0,x0,times,circ_border)

@@

get_trajectory_plot(data_circ_cubic_repulsive,circ_border)
```

@@

data_circ_hanks=foreach (i=1:N,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

@@

```{r}
get_trajectory_plot(data_circ_hanks,circ_border)
```

# Trajectories in a circular domain with islands
